---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
source(".statapath.R")
source("0_packages.R")

load("_data/ess50prep.Rds")
```


# 1. Random intercept model without independent variables

***

## 1.1 ML-estimation {.tabset}

### Stata
```{r, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ess50.dta", clear

xtmixed stfdem || cntry:  // ML is default in Stata
estat icc
```


### R
##### First: Save Model (like Stata's `est store`)
```{r}
multi1 <- lmer(stfdem ~ (1|cntry), data = ess, REML = FALSE) # REML is default in R
```
`1` is R's way of refering to an intercept

##### Then: Inspect model

###### Coefficients (Lower block in Stata output; std.dev no variance)
```{r}
tidy(multi1)
```

###### Model fit (Upper block in Stata output)
```{r}
glance(multi1)
```

###### Regression values for each observation, head shows only first 6 rows
```{r}
augment(multi1) %>% head()
```

###### Shows intraclass correlation
```{r}
icc(multi1) 
```


## 1.2 REML-estimation {.tabset}
### Stata
##### ML is default in stata, REML-estimation can be achieved with option `, reml`
```{r, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ess50.dta", clear
xtmixed stfdem || cntry:, reml
estat icc
```
### R
##### REML is default in R, option needs to be turned off for ML-estimation (`REML = FALSE`)
```{r}
reml <- lmer(stfdem ~ (1|cntry), data = ess)               
ml   <- lmer(stfdem ~ (1|cntry), data = ess, REML = FALSE) 
icc(reml)
icc(ml)    
```
##### -> REML results differ from ML

## 1.3 Plot Data {.tabset}

### Stata
      not available (?)
      
### R {.active}
```{r message=FALSE, warning=FALSE}
sjp.lmer(multi1, type = "fe.slope", vars = c("country"))
```

