---
output:
  html_document:
    code_folding: "show"
    df_print: paged
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
statapath = "/Applications/Stata/StataMP.app/Contents/MacOS/Stata-MP"

source("0_packages.R")

load("../data/ess50prep.Rds")
```

# 1. Random intercept model without independent variables  {.tabset}
## 1.1 ML-estimation {.tabset}

### Models
#### Stata
```{r, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "../data/ess50.dta", clear

xtmixed stfdem || cntry:  // ML is default in Stata
estat icc
```


#### R
##### First: Save Model (like Stata's `est store`)
```{r}
multi1 <- lmer(stfdem ~ (1|cntry), data = ess, REML = FALSE) # REML is default in R
```
`1` is R's way of refering to an intercept

##### Then: Inspect model

###### Coefficients (Lower block in Stata output; std.dev no variance)
```{r}
tidy(multi1)
```

###### Model fit (Upper block in Stata output)
```{r}
glance(multi1)
```

###### Regression values for each observation, head shows only first 6 rows
```{r}
augment(multi1) %>% head()
```

###### Shows intraclass correlation
```{r}
icc(multi1) 
```


## 1.2 REML-estimation {.tabset}

### Models
#### Stata
##### ML is default in stata, REML-estimation can be achieved with option `, reml`
```{r, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "../data/ess50.dta", clear
xtmixed stfdem || cntry:, reml
estat icc
```

#### R
##### REML is default in R, option needs to be turned off for ML-estimation (`REML = FALSE`)
```{r}
reml <- lmer(stfdem ~ (1|cntry), data = ess)               
ml   <- lmer(stfdem ~ (1|cntry), data = ess, REML = FALSE) 
icc(reml)
icc(ml)    
```
##### -> REML results differ from ML

## 1.3 Plot Data

#### R
```{r message=FALSE, warning=FALSE}
sjp.lmer(multi1, type = "fe.slope", vars = c("country"))
```

