---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
source(".statapath.R")
source("0_packages.R")
load("_data/ess50prep.Rds")
```

# 2. Random intercept model with independent variables 

***

## 2.1 Random Intercept Model {.tabset}

### Stata
```{r, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ess50.dta", clear

xtmixed stfdem polintr nwsptot gndr || cntry:
estat icc
```

### R
#### First: Save Model (like Stata's `est store`)
```{r}
multi2 <- lmer(stfdem ~ polintr + nwsptot + gndr + (1|cntry), data = ess, REML = FALSE)
```

#### Then: Inspect model
```{r}
tidy(multi2)
glance(multi2)
icc(multi2)
```

#### R-Squared, works for random intercept models only
```{r}
r2(multi2)
```

#### Predictions for individual cluster
```{r}
ranef(multi2) 
```


## 2.2 Fixed Effects Plots {.tabset}
### Stata
      not (yet) available
      
### R {.active}
#### Fixed effects slopes for each random intercept 
```{r}
sjp.lmer(multi2, type = "ri.slope")
```

#### Fixed effects slopes for each random intercept highlighting **Portugal** and **Denmark**
```{r}
sjp.lmer(multi2, type = "ri.slope", emph.grp = c("PT", "DK"), facet.grid = FALSE)
```
