---
output:  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
source(".statapath.R")
source("0_packages.R")

load("_data/ess50prep.Rds")
```

# 3. Random Coefficient Model 

***

## 3.1 Without Gender Variable   {.tabset}

### Stata
```{r, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ess50prep.dta", clear

xtmixed stfdem polintr nwsptot || cntry: polintr nwsptot, cov(unstructured)
estat icc
```

### R
#### First: Save Model (like Stata's `est store`)
```{r}
multi3 <- lmer(stfdem ~ polintr + nwsptot + (1 + polintr + nwsptot |cntry), data = ess, REML = FALSE)

```
`1 + ` is not necessarily needed because random coefficient models imply that the intercept is also random

#### Then: Inspect model
```{r}
tidy(multi3)
glance(multi3)
icc(multi3)
```

## 3.2 With gender variable    {.tabset}

### Stata
```{r no converge, results = T, engine = "stata", engine.path= statapath, comment = ""}
set maxiter 20
use "_data/ess50prep.dta", clear

xtmixed stfdem || cntry: polintr nwsptot gndr, cov(unstructured)
```

### R
#### First: Save Model (like Stata's `est store`)
```{r}
multi3 <- lmer(stfdem ~ (1 + polintr + nwsptot + gndr |cntry), data = ess, REML = FALSE)
```

#### Then: Inspect model
```{r}
tidy(multi3)
glance(multi3)
icc(multi3)
```

## 3.3 Plotting data {.tabset}

### Stata
not (yet) available

### R {.active}
#### Random effect plots
##### Sorted alphabetically
```{r message=FALSE, warning=FALSE}
sjp.lmer(multi3, y.offset = .6)
```

##### Everything sorted by values of intercept
```{r message=FALSE, warning=FALSE}
sjp.lmer(multi3, sort.est = "(Intercept)", y.offset = .6)
```

##### Every plot sorted by its own values
```{r message=FALSE, warning=FALSE}
sjp.lmer(multi3, sort.est = "sort.all", y.offset = .6, facet.grid = FALSE)
```

#### Random slopes plots
##### Random slopes depending on random intercepts
```{r}
sjp.lmer(multi3, type = "rs.ri", vars = "polintr", show.legend = TRUE)
```

##### Random slopes depending on random intercepts highlighting **Denmark**, **Portugal**, **Luxemburg**
```{r}
sjp.lmer(multi3, type = "rs.ri", vars = "polintr", sample.n = c(6, 16, 20),  show.legend = TRUE)
```
