---
output:
  html_document:
    code_folding: "show"
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
source(".statapath.R")
source("0_packages.R")
options(digits=10)
```

# Prepare Data for Exercises {.tabset}

## Stata
```{r stata create and clean, results = T, engine = "stata", engine.path= statapath, comment = ""}
set more off
capt clear
version 14


* 1. Load datasets and merge (also exercise 1.1)
********************************************************************************
use "_data/pgen.dta", clear

merge 1:1 pid syear using "_data/ppfadl.dta" , ///
      keepusing(sampreg psample pop sex gebjahr phrf phrf0 phrf1) keep(match) nogen

* 2. Data cleaning
********************************************************************************
* 2.1 define missing values
	mvdecode *, mv(-8/-1)
	
* 2.2 generierate variables (also exercise 1.2)
*************************************
** demographics
	cap drop ost
	recode sampreg 1 = 0 2 = 1, gen(ost)
	lab def ost 0 "west" 1 "east" , replace
	lab val ost ost
	tab1 ost sampreg

	cap drop frau
	recode sex 1 = 0 2 = 1, gen(frau)
	lab def frau 0 "male" 1 "femal" , replace
	lab val frau frau
	tab1 frau sex
	
	cap drop alter
	gen alter = syear - gebjahr

** Job Status
	cap drop erwstatus
	recode pgemplst 1 = 1  2 4 = 2 * = ., gen(erwstatus)
	lab def erwstatus 1 "fulltime" 2 "part-time", replace
	lab val erwstatus erwstatus

*** Consumer Price Index Jahr 2010 (also Exercise 1.2)
	cap drop cpi
	gen cpi=.
	label var cpi Preisindex
	replace cpi=(106.9) if syear==2015  // laut destatis
	replace cpi=(106.6) if syear==2014  // laut destatis
	replace cpi=(105.7) if syear==2013  
	replace cpi=(104.1) if syear==2012
	replace cpi=(102.1) if syear==2011
	replace cpi=(100)  if syear==2010   // base year 2010
	replace cpi=(98.9) if syear==2009
	replace cpi=(98.6) if syear==2008
	replace cpi=(96.1) if syear==2007
	replace cpi=(93.9) if syear==2006
	replace cpi=(92.5) if syear==2005   
	replace cpi=(91)   if syear==2004
	replace cpi=(89.6) if syear==2003
	replace cpi=(88.6) if syear==2002
	replace cpi=(87.4) if syear==2001
	replace cpi=(85.7) if syear==2000
	replace cpi=(84.4) if syear==1999 & ost==0
	replace cpi=(85.1) if syear==1999 & ost==1
	replace cpi=(83.8) if syear==1998 & ost==0
	replace cpi=(84.8) if syear==1998 & ost==1
	replace cpi=(83.1) if syear==1997 & ost==0
	replace cpi=(83.9) if syear==1997 & ost==1
	replace cpi=(81.5) if syear==1996 & ost==0
	replace cpi=(82.0) if syear==1996 & ost==1
	replace cpi=(80.5) if syear==1995 & ost==0
	replace cpi=(80.5) if syear==1995 & ost==1
	replace cpi=(79.2) if syear==1994 & ost==0
	replace cpi=(79.0) if syear==1994 & ost==1
	replace cpi=(77.1) if syear==1993 & ost==0
	replace cpi=(76.2) if syear==1993 & ost==1
	replace cpi=(74.5) if syear==1992 & ost==0
	replace cpi=(68.9) if syear==1992 & ost==1
	replace cpi=(71.6) if syear==1991 & ost==0
	replace cpi=(60.8) if syear==1991 & ost==1
	replace cpi=(69.1) if syear==1990 & ost == 0
	replace cpi=(67.3) if syear==1989 & ost == 0
	replace cpi=(65.4) if syear==1988 & ost == 0
	replace cpi=(64.6) if syear==1987 & ost == 0
	replace cpi=(64.5) if syear==1986 & ost == 0
	replace cpi=(64.6) if syear==1985 & ost == 0
	replace cpi=(63.3) if syear==1984 & ost == 0
	

*** Brutto Income in prices from 2010
	cap drop breink10
	gen breink10= pglabgro/cpi*100
	lab var breink10 "Monthly Brutto Income in 2010 Prices"

*** Brutto Income in Prices from 2015
	cap drop breink15
	gen breink15= pglabgro/cpi*106.9
	lab var breink15 "Monthly Brutto Income in 2015 Prices"

*** Hourly Wages in Prices from 2015
	cap drop hwageb
	gen hwageb = breink15/(pgtatzeit*(52/12))
	lab var hwageb "Hourly Brutto Income in 2015 Prices"

*** Work Experience
	cap drop erf
	gen erf= pgexpft + .4*pgexppt
	gen erfq= erf^2

*** Labor Market Segments
	cap drop qualstand
	gen qualstand= pgausb
	recode qualstand 1/4=0 5/9=1

	cap drop intern
	gen intern = pgallbet
	recode intern 1/2=0 3/5=1 
	tab intern qualstand
	
*** Centralize years of schooling and work experience
	gen cpgbilzeit = pgbilzeit - r(mean)
	gen cerf = erf -r(mean)
	
*** Log of Wage
	gen lnwage= ln(hwageb)
	lab var lnwage "Log. Hourly Brutto Wages in 2015 Prices"


* 4. Define Sample for Analysis
********************************************************************************
      cap drop asample
      gen asample=1 if 	pgtatzeit >= 6 & !missing(pgtatzeit) & ///
						alter <= 65 & alter >= 18 & !missing(alter)	&	///
						(pgemplst==1 | pgemplst==2 | pgemplst==4) &  ///
						pop<3
					 
* unplausible cases:
	replace asample=. if    pid==1380202 | ///
                  	      (pid== 8267202 & syear==2007) | ///
                  	      (pid==2633801 & syear==2006) | ///
                  	      (pid==2582901 & syear > 2006) | ///
                  	      pid==1380202 | ///
                  	      pid == 607602  | ///
                  	      pid==2555301
      
*** Save Analysis Sample Data
      save "_data/ex_mydf.dta", replace
```

## R {.tabset}

### Read SOEP Data into R 
Because SOEP Data are not available in R format we can read them into R with `readstata13` or `haven` and save them as .rds for future use.
```{r create pgen ppfadl}
#### 1. generate  and merge dataset mydf_raw  
# PGEN
  if (file.exists(paste0("_data/pgen.rds"))) {
        pgen <- readRDS("_data/pgen.rds")
  } else{
        cat("RDS Datei nicht vorhanden, wird aus Stata Dataset .dta erstellt")
        pgen <- read_dta("_data/pgen.dta")
        saveRDS(pgen, file = "_data/pgen.rds")
  }

# PPFADL
  if (file.exists(paste0("_data/ppfadl.rds"))) {
        ppfadl <- readRDS("_data/ppfadl.rds")
  } else{
        cat("RDS Datei nicht vorhanden, wird aus Stata Dataset .dta erstellt")
        ppfadl <- read_dta("_data/ppfadl.dta") %>% 
            tbl_df() %>% 
            select(sampreg,
                   psample,
                   pop,
                   sex,
                   gebjahr,
                   phrf,
                   phrf0,
                   pid,
                   syear,
                   phrf1)
        saveRDS(ppfadl, file = "_data/ppfadl.rds")
  }
```

### Get Additional Data
#### CPI (base 2010)
Note: **Separate Calculation of CPI in SOEP from 1991 to 2000** the SOEP provides separate cpi for former East and West- Germany from the Years 1991 to 2000. These are reflected in the cpi_s (_s for SOEP) variable. One thing to remark is that the SOEP does not provide CPI's for East-Incomes in 1990. Those deflated incomes can therefore not be calculated. 
```{r cpi, warning = F}
# CPI (base 2010)
# download.file("http://research.stlouisfed.org/fred2/data/DEUCPIALLMINMEI.txt", "~/Desktop") 

# look for how many lines to skip
cpi_url <- "https://research.stlouisfed.org/fred2/data/DEUCPIALLMINMEI.txt"
cpi_url2 <- "https://www.destatis.de/DE/Publikationen/Thematisch/Preise/Verbraucherpreise/VerbraucherpreisindexLangeReihenXLS_5611103.xlsx"

# cpi from the net
cpi_raw <-read.table(cpi_url, skip = 15, header = TRUE)
cpi <- cpi_raw %>% 
      mutate(cpi_year = year(.$DATE)) %>% 
      # convert from monthly to yearly cpi values:
      group_by(cpi_year) %>% 
      dplyr::summarize(cpi = mean(VALUE)) %>% 
      # create adjustment factor relative 2010 year's prices:
      mutate(adj_fac2010 = cpi/cpi[cpi_year == 2010],
             adj_fac2015 = cpi/cpi[cpi_year == 2015])

# cpi from SOEP
cpi_raw_s <- read.table("_data/cpi.csv", header = T, na.strings = "-", dec = ".",
                        numerals = c("no.loss"))
cpi_s <- cpi_raw_s %>% 
      select(year, cpi_west, cpi_ost) %>%
      # add missing values based on destatis data
      add_row(year = 2014:2015, 
              cpi_ost = c(106.6, 106.9),
              cpi_west = c(106.6, 106.9)) %>% 
      mutate(syear = as.numeric(year)
             ) %>% 
      select(-year)
```

### Create Dataset {.active}
#### 1. Merge Data
```{r merge data r, warning=F}
#  merge datasets
mydf_raw <- ppfadl %>% 
      inner_join(y = pgen, by= c("pid", "syear")) %>% 
      left_join(y = cpi, by = c("syear" = "cpi_year")) %>% 
      left_join(cpi_s, by = c("syear")) %>% 
      mutate(cpi_s = case_when(
            sampreg == 2 ~ cpi_ost, 
            sampreg == 1 ~ cpi_west)) %>% 
      select(-cpi_west, -cpi_ost)
```

#### 2. Clean Data
##### 2.1 Define Missing Codes
```{r define missings}
      na_codes <- c(-1: -8)
      for (i in seq_along(mydf_raw)) {
            mydf_raw[[i]][mydf_raw[[i]] %in% na_codes] <- NA
      }
```

##### 2.2 Generate Variables
Note:

**Logarithm of wage differences in STATA and R**. 

example:
      - R `mutate(data, lnwage = log(wage))`
      - STATA `gen lnwage = ln(wage)`
- In R the function log() calculates the natural logarithm of wage as described [here](https://stackoverflow.com/questions/24304936/r-using-equation-with-natural-logarithm-in-nls), returns -Inf if wage = 0.
- In STATA ln() returns missings for lnwage if wage is 0. 
- If you want the same results, you can code `mutate(data, lnwage = ifelse(wage <0, log(wage), NA)`.

```{r generate variables}
# 2.2 Generate Variables
# Employment Status
mydf_raw$erwstatus <- mydf_raw %>% 
      mutate(erwstatus = case_when(
                              pgemplst == 1 ~ 1, 
                              pgemplst %in% c(2, 4) ~ 2
                              )
            ) %>%
      select(erwstatus) %>% unlist() %>% 
      labelled(c("Vollzeit" = 1, "Teilzeit" = 2)) 


ex_mydf <- mydf_raw %>% 
      # Demographics
      mutate(ost = as_factor(labelled(sampreg - 1, c("west" = 0, "east" = 1))), 
             # age
             alter = syear - gebjahr, 
             # gender
             frau = as_factor(labelled(sex - 1, c("male" = 0, "female" = 1))), 
      # Generate Brutto Income 2010 and 2015, Hourly Wage 2015
             # Einkommen, Hourly Wage
             breink10 = pglabgro / cpi_s * 100,
             # deflate income for 2015
             breink15 = pglabgro / cpi_s * 106.9,
             # hourly wage in prices of 2015
             hwageb = breink15/(pgtatzeit*(52/12)),
             # Log of wage
             lnwage = ifelse(hwageb > 0, log(hwageb), NA),
             # job experience
             erf = pgexpft + 0.4*pgexppt,
             # Arbeitsmarktsegmente I
             qualstand = case_when(
                   between(pgausb, 1,4) ~ 0,
                   between(pgausb, 5,9) ~ 1),
             # Arbeitsmarktsegments II
             intern = case_when(
                   between(pgallbet, 1,2) ~ 0,
                   between(pgallbet, 3,5) ~ 1),
            # create id var
            pid_syear = (paste(pid, syear, sep = "." ))
                  ) %>% 
      group_by(syear) %>% 
      mutate( 
            # centralized values for education by year
            cpgbilzeit = pgbilzeit - mean(pgbilzeit, na.rm = T),
            cerf = erf - mean(erf, na.rm = T)
            ) %>% 
      ungroup()
```

##### 2.3 save dataset
```{r save data}
saveRDS(ex_mydf, file = "_data/ex_mydf.rds")
```

