---
output:
  html_document:
    code_folding: "show"
    df_print: paged
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
source("0_packages.R")
source(".statapath.R")
knitr::opts_chunk$set(echo = TRUE, cache = T)

```

# Exercise 1 {.tabset}

## Questions
1.  Use the long-file pgen as master data und 
    merge the variables sampreg, psample, pop, sex, gebjahr and phrf 
    from the file ppfadl to it. You will find gross labor income in the variable pglabgro.

2.  You will have to transform the income data using a consumer price index
    (see table A2 in SOEPmonitor).
    
3.  Replicate the tables concerning  current monthly individual gross labor income from SOEP-Monitor (p.73): https://www.diw.de/de/diw_02.c.222728.de/soepmonitor.html

## Data Prep {.tabset}

### Stata
```{r stata create and clean, results = F, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ex_mydf.dta", clear
	
* Sample for Analysis
**********************************************
	tab pop

	cap drop asample
	gen asample=1 if pop<3 				// only private households
```

### R 
```{r load data }
ex_mydf <- readRDS(file = "_data/ex_mydf.rds")

asample <- ex_mydf %>% filter(pop <3)
```

## Answers {.active}
### 1.1 Create Data
**Use the long-file pgen as master data und merge the variables sampreg, psample, pop, sex, gebjahr and phrf from the file ppfadl to it. You will find gross labor income in the variable pglabgro.**
    
See [data creation](ex0_create.html)

### 1.2 Wrangle Data
**You will have to transform the income data using a consumer price index (see table A2 in SOEPmonitor).**

Also see [data creation](ex0_create.html)

### 1.3 Replicate Table {.tabset}
**Replicate the tables concerning  current monthly individual gross labor income from SOEP-Monitor (p.73): https://www.diw.de/de/diw_02.c.222728.de/soepmonitor.htmlDescribe Data or download [here](info/diw_monitor_individuals_2013.pdf#page=73)**

#### Stata
```{r stata 1.3 describe, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ex_mydf.dta", clear
	
* Sample for Analysis
**********************************************
	tab pop

	cap drop asample
	gen asample=1 if pop<3 				// Anstaltsbevölkerung rausnehmen
	
* 3. Replication of SOEP-Monitor (Exercise 1)
***************************************************************************
	table syear sampreg   [pw=phrf], c(mean breink10) center col
	table syear erwstatus [pw=phrf], c(mean breink10) center col by(sampreg) format(%9.2fc)
```

#### R {.active}
```{r 1.3 describe}
#### 1.3 Describe Data ####
# # Means by Region 
# prep1.3 <- asample %>% 
#       select(syear, breink10, erwstatus, ost, phrf) %>% 
#       dplyr::group_by(syear, ost) %>% 
#       # mutate_at(c("breink10"), funs(Total = mean(., na.rm = T))) %>% 
#       dplyr::summarise(TotalMean = mean(breink10, weights = phrf, na.rm = T))
# 
#      xtabs(TotalMean ~ syear + ost, prep1.3) 

# Means by Region and Job Status
result1.3 <- asample %>% 
      group_by(syear, ost, erwstatus) %>%
      dplyr::summarise(mean = mean(breink10, weights = phrf, na.rm = T))

xtabs(mean ~ syear + erwstatus + ost, result1.3) 
```




