---
output: html_document
---

```{r setup, include=FALSE}
source("0_packages.R")
source(".statapath.R")

knitr::opts_chunk$set(echo = TRUE, cache = T)

```

# Exercise 4 {.tabset}
## Questions

4.1 Estimate the Mincer equation of last weeks exercise seperately for East and West Germany for all years since 1992 using loops (see foreach).

4.2 Present the effects of years of education and of sex (the gender pay gap) graphically showing the trend since 1992 (see collapse) Tip: Use the Stata commands foreach, collapse and graph combine.

4.3 Tell Stata your data is panel data using xtset. Describe the variables of your Mincer equation and differentiate within and between variation (xtsum).


## Data Prep {.tabset}
### Stata
```{r stata create and clean, results = T, engine = "stata", engine.path= statapath, comment = ""}
set more off
capt clear
version 14

use "_data/ex_mydf.dta", clear
```
### R
```{r load data}
#### load dataset ####
ex_mydf <- readRDS(file = "_data/ex_mydf.rds")

asample <- ex_mydf %>% 
      filter(
            # Working Hours
            pgtatzeit >= 6,
            # age
            alter %>% dplyr::between(18, 65),
            # Employment Status
            pgemplst %in% c(1,2,4),   # full-time(1), part-time(2), marg.-empl(4)
            # population status
            pop < 3 # only private households
            ) %>%
      # filter unplausible cases
      mutate(na = case_when(
              pid %in% c(1380202, 1380202, 607602, 2555301) ~ 1,
              pid == 8267202 & syear == 2007 ~ 1,
              pid == 2633801 & syear == 2006 ~ 1,
              pid == 2582901 & syear > 2006 ~ 1 )
             ) %>% 
      filter(is.na(na)) %>% 
      # select relevant variables
      dplyr::select(hwageb, lnwage, pgbilzeit, pgisco88, pgnace, pgexpue,
             sex, ost, erf, frau, pgallbet, phrf, syear, pid, pid_syear) 
      # %>%
      # drop_na() 

```
## Answers {.active}

### 4.1 Mincer Equation {.tabset}
Estimate the Mincer equation of last weeks exercise seperately for East and West Germany for all years since 1992 using loops (see foreach).

#### Stata
```{r stata 4.1, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ex_mydf.dta", clear

cap drop educationrev 
gen educationrev=.
cap drop upper lower
gen upper=.
gen lower=.

foreach year of numlist 1991/2015 {
      qui: reg lnwage pgbilzeit c.erf##c.erf pgexpue ost frau i.pgallbet if asample==1 & syear==`year' [pw=phrf], eform(b) cluster(hid)
      replace educationrev = exp(_b[pgbilzeit]) if syear==`year'
      replace upper = exp(_b[pgbilzeit]+ 1.96*_se[pgbilzeit]) if syear==`year'
      replace lower = exp(_b[pgbilzeit]- 1.96*_se[pgbilzeit]) if syear==`year'
}

preserve
collapse educationrev upper lower, by(syear) 
twoway (connected educationrev syear, sort) ///
(line upper lower syear, sort lpattern(dash dash) lcolor(bluishgray8 bluishgray8)) ///
, legend(off) subtitle(Gesamt)

graph save Graph "out/educationrev_gesamt.gph", replace
restore

** repeat for only West Germany

cap drop educationrev_west
gen educationrev_west =.
cap drop upper lower
gen upper=.
gen lower=.
foreach year of numlist 1984/2015 {
reg lnwage pgbilzeit c.erf##c.erf pgexpue ost frau i.pgallbet if asample==1 & syear==`year' & ost == 0 [pw=phrf], eform(b) cluster(hid)
replace educationrev_west = exp(_b[pgbilzeit]) if syear==`year'
replace upper = exp(_b[pgbilzeit]+ 1.96*_se[pgbilzeit]) if syear==`year'
replace lower = exp(_b[pgbilzeit]- 1.96*_se[pgbilzeit]) if syear==`year'
}


preserve
collapse educationrev_west upper lower, by(syear) 
twoway (connected educationrev_west syear, sort) ///
(line upper lower syear, sort lpattern(dash dash) lcolor(bluishgray8 bluishgray8)) ///
, legend(off) subtitle(West Germany)
graph save Graph "out/educationrev_west.gph", replace
restore


** repeat for only East Germany

cap drop educationrev_ost
gen educationrev_ost =.
cap drop upper lower
gen upper=.
gen lower=.
foreach year of numlist 1991/2015 {
reg lnwage pgbilzeit c.erf##c.erf pgexpue ost frau i.pgallbet   if asample==1 & syear==`year' & ost == 1 [pw=phrf], eform(b) cluster(hid)
replace educationrev_ost = exp(_b[pgbilzeit]) if syear==`year'
replace upper = exp(_b[pgbilzeit]+ 1.96*_se[pgbilzeit]) if syear==`year'
replace lower = exp(_b[pgbilzeit]- 1.96*_se[pgbilzeit]) if syear==`year'
}

preserve
collapse educationrev_ost upper lower, by(syear) 
twoway (connected educationrev_ost syear, sort) ///
(line upper lower syear, sort lpattern(dash dash) lcolor(bluishgray8 bluishgray8)) ///
, legend(off) subtitle(East Germany)
graph save Graph "out/educationrev_ost.gph", replace
restore

***
graph combine  ///
      "out/educationrev_west.gph"  ///
      "out/educationrev_ost.gph"  ///
      "out/educationrev_gesamt.gph" ///
, ycommon xcommon note(GSOEP 32, size(vsmall) position(5)) title(Development of Education Revenues)

graph save Graph "out/educationrev_all.gph", replace
```
 
#### R {.tabset .active}
##### Model
Trend by Year
```{r model 4.1 by year}
# Fit Hourly Wage Brutto Model by year
      fit4.1 <- asample %>% 
            group_by(syear) %>% 
            do(model = lm(lnwage ~ 0+ pgbilzeit + I(erf^2) + pgexpue +
                                ost + frau + factor(pgallbet), 
               data=., weights=phrf))
```

Trend by Year and Region
```{r model 4.1 by region and year}
fit4.1_region <- asample %>% 
      group_by(ost, syear) %>% 
      do(model = lm(lnwage ~ 0 + pgbilzeit + I(erf^2) + pgexpue +
                          ost + frau + factor(pgallbet), 
         data=., weights=phrf))

      # look at model
            # (fit4_region$model) # only use if not too many models
            #fit4_region %>% tidy(model)
            #fit4_region %>% glance(model)
            #fit4_region %>% augment(model)
```

##### Plot {.active}
Education Revenue Trend 1992 - 2015
```{r 4.1 Education Revenue Trend}
     # grab relevant info
      td_fit <- fit4.1 %>% tidy(model, conf.int = TRUE)
      td_plot <- td_fit %>% filter(term == "pgbilzeit",
                                   syear > 1990) 
      
      # Plot Over Years
      td_plot %>% 
            ggplot(aes(x = syear, y = estimate)) +
            geom_point() +
            geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                          alpha=0.3, color = "black") +
            geom_line() +
            ylab("Education Revenue") +
            xlab("Survey Year") +
            ggtitle("Education Revenue from 1991 to 2015") +
            scale_x_continuous(breaks = seq(1991, 2016, by= 5))+
            scale_y_continuous(labels = scales::percent)+
            theme_bw()
```

Education Revenue Trend over Time and Region
```{r 4.1 Education Revenue Trend by Region}
      # make tidy dataframe 
      td_fit <- fit4.1_region %>% tidy(model, conf.int = TRUE)
      # select relevant info (coefficient for gender)
      td_plot <- td_fit %>% filter(term == "pgbilzeit")
      
      # plot Trend by Region (grouped)
      td_plot %>% 
      ggplot(aes(x = syear, y = estimate, 
                 color = as.factor(ost)), fill = as.factor(ost)) +
            geom_point() +
            geom_ribbon(ggplot2::aes_string(ymin = "conf.low", ymax = "conf.high", 
                fill = "as.factor(ost)", color = NULL), alpha = 0.2) +
            geom_line() +
            ylab("Education Revenue") +
            xlab("Survey Year") +
            ggtitle("Education Revenue from 1984 to 2015 by Region") +
            scale_x_continuous(breaks = seq(1985, 2015, by= 5))+
            scale_y_continuous(labels = scales::percent)
     
      
      # plot Trend by Region (facet)
      td_plot %>% 
      ggplot(aes(x = syear, y = estimate)) +
            geom_point() +
            geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                          alpha=0.3, color = "black") +
            geom_line() +
            facet_wrap(~ ost)+
            ylab("Education Revenue") +
            xlab("Survey Year") +
            ggtitle("Education Revenue from 1984 to 2015 by Region") +
            scale_x_continuous(breaks = seq(1985, 2015, by= 5))+
            scale_y_continuous(labels = scales::percent)
```

### 4.2 {.tabset }
Present the effects of years of education and of sex (the gender pay gap) graphically showing the trend since 1992 (see collapse) Tip: Use the Stata commands foreach, collapse and graph combine. 

#### Stata
```{r stata 4.2 Gender Revenue, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ex_mydf.dta", clear

cap drop genderrev
gen genderrev =.
cap drop upper lower
gen upper=.
gen lower=.

foreach year of numlist 1991/2015 {
      reg lnwage pgbilzeit c.erf##c.erf pgexpue ost frau i.pgallbet if asample==1 & syear==`year' [pw=phrf], eform(b) cluster(hid)
      replace genderrev = exp(_b[frau]) if syear==`year'
      replace upper = exp(_b[frau]+ 1.96*_se[frau]) if syear==`year'
      replace lower = exp(_b[frau]- 1.96*_se[frau]) if syear==`year'
}


preserve
collapse genderrev upper lower, by(syear) 
twoway (connected genderrev syear, sort) ///
(line upper lower syear, sort lpattern(dash dash) lcolor(bluishgray8 bluishgray8)) ///
, legend(off) subtitle(Gesamt)
graph save Graph "out/genderrev_gesamt.gph", replace
restore

** repeat for only west germany
cap drop genderrev_west
gen genderrev_west=.
cap drop upper lower
gen upper=.
gen lower=.
foreach year of numlist 1984/2015 {
reg lnwage frau c.erf##c.erf pgexpue ost frau i.pgallbet   if asample==1 & syear==`year' & ost==0 [pw=phrf], eform(b) cluster(hid)
replace genderrev_west = exp(_b[frau]) if syear==`year'
replace upper = exp(_b[frau]+ 1.96*_se[frau]) if syear==`year'
replace lower = exp(_b[frau]- 1.96*_se[frau]) if syear==`year'
}
*twoway (connected genderrev_west syear, sort) (line upper lower syear, sort) 
preserve
collapse genderrev_west upper lower, by(syear) 
twoway (connected genderrev_west syear, sort) ///
(line upper lower syear, sort lpattern(dash dash) lcolor(bluishgray8 bluishgray8)) ///
, legend(off) subtitle(Ostdeutschland)
graph save Graph "out/genderrev_west.gph", replace
restore


** repeat for only east germany

cap drop genderrev_ost
gen genderrev_ost=.
cap drop upper lower
gen upper=.
gen lower=.
foreach year of numlist 1991/2015 {
reg lnwage frau c.erf##c.erf pgexpue ost frau i.pgallbet   if asample==1 & syear==`year' & ost==1 [pw=phrf], eform(b) cluster(hid)
replace genderrev_ost = exp(_b[frau]) if syear==`year'
replace upper = exp(_b[frau]+ 1.96*_se[frau]) if syear==`year'
replace lower = exp(_b[frau]- 1.96*_se[frau]) if syear==`year'
}
*twoway (connected genderrev_ost syear, sort) (line upper lower syear, sort)
preserve
collapse genderrev_ost upper lower, by(syear) 
twoway (connected genderrev_ost syear, sort) ///
(line upper lower syear, sort lpattern(dash dash) lcolor(bluishgray8 bluishgray8)) ///
, legend(off) subtitle(Westdeutschland)
graph save Graph "out/genderrev_ost.gph", replace
restore

***
graph combine  ///
"out/genderrev_ost.gph"  ///
"out/genderrev_west.gph"  ///
"out/genderrev_gesamt.gph" ///
, ycommon xcommon note(GSOEP 32, size(vsmall) position(5)) title(Development of Gender Revenues)

graph save Graph "out/genderrev_all.gph", replace

```
#### R {.tabset .active}
##### Model
The Plots are based on the same models as in 4.1.

##### Plot {.active}
Gender Revenue Trend 1991 - 2015
```{r 4.2 Education Revenue Trend}
     # grab relevant info
      td_fit <- fit4.1 %>% tidy(model, conf.int = TRUE)
      td_plot <- td_fit %>% filter(term == "frau",
                                   syear > 1990) 
      
      # Plot Over Years
      td_plot %>% 
            ggplot(aes(x = syear, y = estimate)) +
            geom_point() +
            geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                          alpha=0.3, color = "black") +
            geom_line() +
            ylab("Gender Revenue") +
            xlab("Survey Year") +
            ggtitle("Gender Pay Gap from 1991 to 2015") +
            scale_x_continuous(breaks = seq(1991, 2016, by= 5))+
            scale_y_reverse(labels = scales::percent)+
            theme_bw()
```

Gender Revenue Trend by Year and Gender
```{r 4.2 Gender Revenue Trend by Region}
      # make tidy dataframe 
      td_fit <- fit4.1_region %>% tidy(model, conf.int = TRUE)
      # select relevant info (coefficient for gender)
      td_plot <- td_fit %>% filter(term == "frau")
      
      # plot Trend by Region (grouped)
      td_plot %>% 
      ggplot(aes(x = syear, y = estimate, color = as.factor(ost))) +
            geom_point() +
            geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                          alpha=0.3, color = "black") +
            geom_line() +
            ylab("Gender Revenue") +
            xlab("Survey Year") +
            ggtitle("Gender Pay Gap by Region") +
            scale_x_continuous(breaks = seq(1985, 2015, by= 5))+
            scale_y_reverse(labels = scales::percent)
     
      # plot Trend by Region (facet)
      td_plot %>% 
      ggplot(aes(x = syear, y = estimate)) +
            geom_point() +
            geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                          alpha=0.3, color = "black") +
            geom_line() +
            facet_wrap(~ost)+
            ylab("Gender Revenue") +
            xlab("Survey Year") +
            ggtitle("Gender Pay Gap by Region") +
            scale_x_continuous(breaks = seq(1985, 2015, by= 5))
```

### 4.3 xtset {.tabset}
Tell Stata your data is panel data using xtset. Describe the variables of your Mincer equation and differentiate within and between variation (xtsum).

***

* within: bei Min und Max muss noch der Grand-Mean abgezogen werden, 
* um die Abweichung von den personenspezifischen Mittelwerten zu erhalten

#### Stata
```{r stata 4.3 xtsum, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ex_mydf.dta", clear

tab pop
tab pgstib

*Only private Households (pop > 3), not in wf (10), apprentice (11), unempl (12), retired (13)
cap drop asample
gen asample = 1 if pop < 3 & pgstib!=10 & pgstib!=11 & pgstib!=12 & pgstib != 13  
	
xtset pid syear
xtdes

xtreg lnwage c.erf##c.erf pgexpue frau i.pgallbet if asample==1, fe

* xtsum
xtsum lnwage pgbilzeit erf ost frau if asample==1

*** Tabellen
* xttab
xtab erf
xtab pgexpue
xtab frau
xtab pgallbet
xtab pgbilzeit

```

#### R
In R there is no package available that calculated the between and within variance of variablen in panel datasets. Therefore we need to write a function to achieve the desired goal. The function used in this case originates from [here](https://stackoverflow.com/questions/49282083/xtsum-command-for-r) and has been adapted a bit.

In a first step the overall descriptive statistics are calculated and stored,
then the between means are calculated for each individual and then the variation and other statistics are calculated. Lastly the stats for within variation are calculated and the stored results are returned. The function can now be used in the following way:

The function XTSUM takes three inputs:

data -- the dataset
varname -- the variable to xtsum
unit -- the identifier for the within dimension 

```{r}
asample <- ex_mydf %>% 
      filter(
            # only private households
            pop <= 2,
            # only workforce
            # not in wf (10), apprentice (11), unempl (12), retired (13)
            !pgstib %in% c(10,11,12,13) 
      )

# quick frequency table
# asample %>% group_by(syear) %>% summarise(n = n()) %>% as.data.frame()

XTSUM <- function(data, varname, unit) {
  varname <- enquo(varname)
  loc.unit <- enquo(unit)

ores <- data %>% summarise(ovr.mean = mean(!! varname, na.rm=TRUE),
                           ovr.sd = sd(!! varname, na.rm=TRUE), 
                           ovr.min = min(!! varname, na.rm=TRUE), 
                           ovr.max = max(!! varname, na.rm=TRUE),
                           ovr.N = sum(as.numeric((!is.na(!! varname)))),
                           ovr.N2 = n())
bmeans <- data %>% 
      group_by(!! loc.unit) %>% 
      summarise(meanx = mean(!! varname, na.rm=T), 
                t.count = sum(as.numeric(!is.na(!! varname))),
                t.count2 = na.omit(n())
                )
                

bres <- bmeans %>% 
      ungroup() %>% 
      summarise(between.sd = sd(meanx, na.rm=TRUE), 
                between.min = min(meanx, na.rm=TRUE), 
                between.max = max(meanx, na.rm=TRUE),
                Units = sum(as.numeric(!is.na(t.count))), 
                t.bar = mean(t.count, na.rm=TRUE),
                t.bar2 = mean(t.count2, na.rm = T))

wdat <- data %>% 
      group_by(!! loc.unit) %>% 
      mutate(W.x = scale(!! varname, scale=FALSE))

wres <- wdat %>% 
      ungroup() %>% 
      summarise(within.sd = sd(W.x, na.rm=TRUE), 
                within.min = min(W.x, na.rm=TRUE), 
                within.max = max(W.x, na.rm=TRUE))

return(list(ores = ores, bres = bres, wres = wres))
}
```

Using the function
```{r xtsum}
XTSUM(asample, varname = lnwage , unit = pid)
XTSUM(asample, varname = pgbilzeit , unit = pid)
XTSUM(asample, varname = frau , unit = pid)
XTSUM(asample, varname = pgallbet , unit = pid)
XTSUM(asample, varname = pgexpue , unit = pid)
```

Links:
- https://stackoverflow.com/questions/14165752/between-within-standard-deviations-in-r
- https://www.r-bloggers.com/dplyr-do-some-tips-for-using-and-programming/
- psdata: https://github.com/rOpenGov/psData/issues/5
- dplyr xtsum: https://github.com/hadley/vctrs/issues/17
- https://christophergandrud.github.io/RepResR-RStudio/
- Get the column number in R given the column name
      - which( colnames(df)=="b" )
- https://stackoverflow.com/questions/16367436/compute-mean-and-standard-deviation-by-group-for-multiple-variables-in-a-data-fr

