---
output:
  html_document:
    code_folding: hide
---


```{r setup, include=FALSE}
source("0_packages.R")
source(".statapath.R")

knitr::opts_chunk$set(echo = TRUE, cache = T)

```

# Exercise 7 {.tabset}
## Questions

7.1 Specify a random and a fixed effects model. Compare the results.

7.2 Execute the hausman test. Which model specification is more appropriate and why?

## Data Prep {.tabset}
### Stata
```{r stata create and clean, results = T, engine = "stata", engine.path= statapath, comment = ""}

set more off
capt clear
version 14

use "_data/ex_mydf.dta", clear
```
### R
#### Load Data
```{r load data}
#### load dataset ####
ex_mydf <- readRDS(file = "_data/ex_mydf.rds")

asample <- ex_mydf %>% 
      filter(
            # Working Hours
            pgtatzeit >= 6,
            # age
            alter %>% dplyr::between(18, 65),
            # Employment Status
            pgemplst %in% c(1,2,4),   # full-time(1), part-time(2), marg.-empl(4)
            # population status
            pop <= 2, # only private households
            # restrict number of years
            syear > 2002) %>%
      # filter unplausible cases
      mutate(na = case_when(
              pid %in% c(1380202, 1380202, 607602, 2555301) ~ 1,
              pid == 8267202 & syear == 2007 ~ 1,
              pid == 2633801 & syear == 2006 ~ 1,
              pid == 2582901 & syear > 2006 ~ 1 )
             ) %>% 
      filter(is.na(na))

# Prepare Panel Dataset
p.asample <- pdata.frame(asample, index = c("pid", "syear"))

# information if panel dataframe
pdim(p.asample)
# summary(p.asample)
```
## Answers {.active}
### 7.1 {.tabset}
Specify a random and a fixed effects model. Compare the results.

***

#### Stata
```{r, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ex_mydf.dta", clear

xtset pid syear
* FE model

xtreg lnwage pgbilzeit c.erf##c.erf pgexpue ost frau i.pgallbet i.syear if asample==1 & syear>2002 ,   fe

* RE model 
xtreg lnwage pgbilzeit c.erf##c.erf pgexpue ost frau i.pgallbet i.syear if asample==1 & syear>2002 ,  re 

```
 
#### R {.tabset .active}
##### Model
```{r model 7.1}

# FE Model
fixed   <- p.asample %>% 
      filter(syear > 2002) %>% 
      plm(lnwage ~  pgbilzeit + erf + I(erf^2) + pgexpue + ost + pgallbet + syear,
             data = p.asample[], 
             model = "within" 
             # weights = phrf
             )

tidy(fixed)

# RE Model
random  <- plm(lnwage ~  pgbilzeit + erf + I(erf^2) + pgexpue + ost + pgallbet + syear,
             data = p.asample, 
             model = "random" 
             # weights = phrf
             )

tidy(random)

```

##### Plot

### 7.2 {.tabset}
Execute the hausman test. Which model specification is more appropriate and why?

***

#### Stata
```{r, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/ex_mydf.dta", clear

xtreg lnwage pgbilzeit c.erf##c.erf pgexpue ost frau i.pgallbet i.syear if asample==1 & syear>2002 ,   fe
estimates store fixed

xtreg lnwage pgbilzeit c.erf##c.erf pgexpue ost frau i.pgallbet i.syear if asample==1 & syear>2002 ,  re 
estimates store random

hausman fixed random
```
 
#### R {.tabset .active}
##### Model
```{r model 7.2}
# Hausman Test
hausman <- phtest(fixed, random)
hausman
```

##### Plot

other things
```{r}
pvar(p.asample)
```

