---
output:
  html_document:
    code_folding: hide
---


```{r setup, include=FALSE}
source("0_packages.R")
source(".statapath.R")

knitr::opts_chunk$set(echo = TRUE, cache = T)
```

# Exercise 8 {.tabset}
## Questions

8.1 Does a divorce (famstat==4) causally affect mental health (mcs)? Construct binary variables for the categories of famstat (tab famstat, gen(fam)) and a binary indicator of low mental health (gen mcs01=mcs<40 if mcs !=.) Estimate the effect of a divorce (fam4) on mental health (mcs01) in two different model specifications: Pooled Logit (logit, cluster()) and FE-Logit (xtlogit, fe). Add relevant third variables, if necessary. Summarize the estimates in a table (est tab)

8.2 Interpret your findings and the differences between the models.

## Data Prep {.tabset}
### Stata
```{r stata create and clean, results = T, engine = "stata", engine.path= statapath, comment = ""}
set more off
capt clear
version 14

* Open the healtl-file.
use "_data/healthl.dta", clear

* Does a divorce (famstat==4) causally affect mental health (mcs)?

/* Construct binary variables for the categories of famstat (tab famstat, gen(fam))
and a binary indicator of low mental health (gen mcs01=mcs<40 if mcs !=.) */
	
	tab famstat, gen(fam)
	tab occupation, gen(occ)

	gen mcs01 = mcs<40 if !missing(mcs)
    
save "_data/health.dta", replace
```
### R
#### Load Data
```{r load data}
#### load dataset ####
health_raw <- read_dta("_data/healthl.dta")
```

Construct binary variables for the categories of famstat and a binary indicator of low mental health (mcs01 = 1 if mcs<40, otherwise 0)
```{r}
# Generate Binary Variables
health <- health_raw %>% 
      # mutate mcs
      mutate(mcs01 = ifelse(mcs < 40, 1, 0)) %>% 
      # generate multiple binary vars from categorical var
      tibble::rownames_to_column() %>%
      group_by_all()%>% 
      dplyr:: summarize(count = n()) %>% 
      spread(famstat, count, sep = "_", fill = 0) %>% 
      ungroup() %>% 
      dplyr:: select(-rowname) %>% 
      drop_na()


health %>% 
      dplyr:: select(mcs, mcs01, starts_with("famstat")) %>% 
      head()

# transform into panel data frame
p.health <- pdata.frame(health, index = c("id", "year"))

```

## Answers {.active}
### 8.1 {.tabset}
Does a divorce (famstat==4) causally affect mental health (mcs)? Construct binary variables for the categories of famstat (tab famstat, gen(fam)) and a binary indicator of low mental health (gen mcs01=mcs<40 if mcs !=.) 

- Estimate the effect of a divorce (fam4) on mental health (mcs01) in two different model specifications: 
      - Pooled Logit (logit, cluster()) and 
      - FE-Logit (xtlogit, fe).
- Add relevant third variables, if necessary. 
- Summarize the estimates in a table (est tab)

***

#### Stata
```{r, results = T, engine = "stata", engine.path= statapath, comment = ""}
use "_data/health.dta", clear

/* Estimate the effect of a divorce (fam4) on mental health (mcs01) in two different model specifications: Pooled Logit (logit, cluster()) and FE-Logit (xtlogit, fe).*/
    
* Pooled Logit
logit mcs01 fam4, cluster(id)
	
* Fixed-Logit 
xtset id year
xtlogit mcs01 fam4, fe
	
* Add relevant third variables, if necessary.

* Pooled Logit
logit mcs01 fam4, cluster(id)

* Fixed-Logit 
xtlogit mcs01 fam4 partner##i.gender i.occupation hhinc, fe

	
* Summarize the estimates in a table (est tab)


* Interpret your findings and the differences between the models.

```
 
#### R {.tabset .active}
##### Models
Note: pglm package is not supported for paneldata:
https://stats.stackexchange.com/questions/146434/why-pglm-fails-for-within-model

Pooled Logit Model
```{r model 8.1}
# Pooled Logit Model
plogit <- glm(mcs01 ~  famstat_4, family = binomial(), data = health)


tidy(plogit)
summary(plogit)
```

Fixed Effects Logit Model
```{r}
# Fixed Effects Logit Model
felogit <- felm(mcs01 ~ famstat_4 | id , data = health)

tidy(felogit)
summary(felogit)

```

- Add relevant third variables, if necessary.
```{r}
# Pooled OLS Model
plogit2 <- glm(mcs01 ~  famstat_4 + partner*gender + factor(occupation)+ hhinc, 
               family = binomial, 
               data = health)

# Fixed Effects Logit Model
felogit2 <- felm(mcs01 ~ famstat_4 + partner + partner:gender + factor(occupation) + hhinc | id , 
                   data = health)

felogit2.2 <- plm(mcs01 ~  famstat_4 + partner*gender + factor(occupation) + hhinc, 
                model = "within", 
                data = p.health)

```

- Summarize the estimates in a table (est tab)
```{r}
summary(plogit2)
summary(felogit2)
summary(felogit2.2)

```

##### Plot

### 8.2 {.tabset}
Interpret your findings and the differences between the models.

***
 
#### R {.tabset .active}
##### Model
```{r model 8.2}

```

##### Plot
