---
author: "Lisa Reiber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook:
    code_folding: show
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Packages and Data
```{r, results=F, message=F, echo=FALSE}
#### install missing packages and load needed packages ####
if (!require("pacman")) install.packages("pacman")
pacman::p_load(foreign, readstata13, ggplot2, lme4, broom, sjPlot, tidyverse)      

#### load dataset ####
ess50 <- read.dta13(file = "../../data/ess50.dta", convert.factor = F)
```


# 1. Hybrid Model
Specify a hybrid model decomposing covariates in between and within effects. Your model should at least include female, new and one additional covariate.

**Data Wrangling**
```{r}
sample <- ess50 %>% 
      mutate(cntry.id = as.numeric(as.factor(cntry)),
             new = ifelse(cntry.id %in% c(7,20,11),1,0),
             new2 = ifelse(cntry.id %in% c(4,12,19,22), 1, 0),
             old = ifelse(new == 0 & new2 == 1 , 1, 0)
             )

# prep hybrid model
sample1 <- sample %>%
  filter(!is.na(polintr), !is.na(nwsptot)) %>% 
  group_by(cntry.id) %>% 
  mutate(m.polintr = mean(polintr), 
         m.nwsptot = mean(nwsptot),  
         m.rel     = mean(rlgdgr),
         dev.polintr = polintr - m.polintr, 
         dev.nwsptot = nwsptot - m.nwsptot,
         dev.rel   = rlgdgr - m.rel)
```

**Hybrid Model**
```{r}
multi <- lmer(stfdem ~ new + new2 + female + m.polintr + dev.polintr +
              m.nwsptot + dev.nwsptot + (1|cntry), data =sample1)

tidy(multi, conf.int = TRUE)

sjp.lmer(multi, y.offset = .6, sort.est = "sort.all", facet.grid = FALSE)
```


# 2. Empirical Bayes Prediction
Calculate Empirical Bayes Predictions for the random intercepts. Present the BLUPS graphically.

```{r}
# ML-Estimates:
ranef(multi)

# Bayes-Predictions:
psi   <- tidy(multi)$estimate[9]
theta <- tidy(multi)$estimate[10]
ess50 %>% 
  group_by(cntry) %>% 
  summarise(clusterN = n()) ->
essN

b.correction <- psi/(psi + theta/essN$clusterN)
b.correction

b.prediction <- cbind(ml = unlist(ranef(multi)), b.correction)
row.names(b.prediction) <- unique(ess50$cntry)

b.prediction %>% 
  as.data.frame() %>% 
  mutate(b.prediction = ml * b.correction)


# varCluster = psi
# varIndi = Theta
# J = cluster

```

# 3. Random Coefficients
Specify at least one level-1-covariate as random coefficient.

```{r}
multi <- lmer(stfdem ~ new + new2 + female + polintr + 
              nwsptot + (1 + polintr + nwsptot|cntry), data = sample)

```

# 4. Empirical Bayes Prediction
Calculate Empirical Bayes Predictions for the random coefficient and present the results graphically.

```{r}
sjp.lmer(multi, y.offset = .6, sort.est = "sort.all", facet.grid = FALSE)
sjp.lmer(multi, type = "rs.ri", vars = "nwsptot")
```

# 5. Cross-Level Interaction
Specify one cross-level-interaction-effect and interpret the results.

```{r}

```

