---
title: ""
author: Lisa Reiber
output: 
  html_notebook: 
    code_folding: show
    toc: yes
    toc_float: true
---

# Exercise 12     
- Open the psuinterview.dta file.
- Estimate 3-level (linear) random intercept model (person, household, interviewer) using political interest as the outcome variable.
- Calculate the intraclass-correlation for each higher level: What is the "contribution" of interviewer and household context for the individual reports on the level of political interest?
- Does the interviewer-level improve model fit significantly?

**Load Packages and Data**
```{r}
#options(scipen = 100)
options(digits = 2)

# install missing packages and load needed packages ####
if (!require("pacman")) install.packages("pacman")
pacman::p_load(readstata13, ggplot2, lubridate, broom, effects, 
               pastecs, tidyverse)      

#load dataset 
psuint <- read.dta13("../../data/psuintviewnew.dta", convert.factors = F)
```

**3-Level random intercept model**
brauche ich hier pid oder psuid? Muss die pid unique für die interviewer id sein?
```{r}
# lmer4
# (formula for random terms | unit for which these terms apply)

skimr::skim(psuint)
sample <- psuint %>% 
      mutate(intid = as.factor(intid),
             hhid = as.factor(hhid))

fit.multi3 <- lmer(polint ~ pid + (1 | intid/hhid), data=sample)
tidy(fit.multi3)

fit.multi3 <- lmer(polint ~ psuid + (1 | hhid/intid), data = sample)
tidy(fit.multi3)


## expands to
# lmer(polint ~ pid * tx + 
#                 (pid | intid:hhid) + 
#                 (pid | intid),
#         data=psuint)

# Stata
# xtmixed risk || _all:R.hhid || intid: || psuid: , mle
```
info: 
- http://rpsychologist.com/r-guide-longitudinal-lme-lmer#conditional-growth-model-dropping-intercept-slope-covariance
- https://stats.stackexchange.com/questions/12768/three-level-hierarchical-regression-using-lmer
- http://lme4.r-forge.r-project.org/slides/2009-07-07-Rennes/3LongitudinalD.pdf
- http://data.princeton.edu/pop510/emg.html
- https://jonathantemplin.com/files/mlm/mlm13icpsr/mlm13icpsr_lecture10.pdf


**Intraclass Correlation**
```{r}
library(sjstats)
# icc(model)
icc(fit.multi3)
```
info: https://github.com/strengejacke/sjstats/blob/master/vignettes/mixedmodels-statistics.Rmd

**Improvement through Interviewer level?**

???

