---
output:
  html_document:
    code_folding: "show"
    df_print: paged
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
source("0_packages.R")

knitr::opts_chunk$set(echo = TRUE, cache = T)
statapath = "/Applications/Stata/StataMP.app/Contents/MacOS/Stata-MP"

```

1.  Replicate the tables concerning 
    current monthly individual gross labor income from SOEP-Monitor (p.73):
    https://www.diw.de/de/diw_02.c.222728.de/soepmonitor.html

2.  Use the long-file pgen as master data und 
    merge the variables sampreg, psample, pop, sex, gebjahr and phrf 
    from the file ppfadl to it. 
    You will find gross labor income in the variable pglabgro.

3.  You will have to transform the income data using a consumer price index
    (see table A2 in SOEPmonitor).#### install missing packages and load needed

Load Packages
```{r load packages,results=F, message=F, echo=FALSE}
#### packages ####
# is there a way to show the packages used from this doc?
```

Paths
```{r}
#### Paths ####
      #inpath <- "M:/Gsoep32long/"                    # Uni
      #inpath <- "S:/DATA/soep32_de_l/stata"          # DIW
      inpath <- "/Volumes/gsoep32l/Gsoep32long" # home
```


```{r}
#### 1. generate  and merge dataset mydf_raw  ####
# PGEN
pgen <- read.dta13(paste0(inpath,"/pgen.dta"), convert.factors = F)
saveRDS(pgen, file = "../../data/pgen.rds")

# PPFADL
ppfadl <- read.dta13(paste0(inpath,"/ppfadl.dta"), convert.factors = F) %>% 
            tbl_df() %>% 
            select(sampreg,
                   psample,
                   pop,
                   sex,
                   gebjahr,
                   phrf,
                   phrf0,
                   pid,
                   syear,
                   phrf1)

saveRDS(ppfadl, file = "../../data/ppfadl.rds")
```

# CPI Preisindex bezogen auf das Jahr 2010
```{r}
# CPI Preisindex bezogen auf das Jahr 2010
# download.file("http://research.stlouisfed.org/fred2/data/DEUCPIALLMINMEI.txt", "~/Desktop") 

# look for how many lines to skip
cpi_url <- "https://research.stlouisfed.org/fred2/data/DEUCPIALLMINMEI.txt"
cpi_url2 <- "https://www.destatis.de/DE/Publikationen/Thematisch/Preise/Verbraucherpreise/VerbraucherpreisindexLangeReihenXLS_5611103.xlsx"

# cpi from the net
cpi_raw <-read.table(cpi_url, skip = 15, header = TRUE)
cpi <- cpi_raw %>% 
      mutate(cpi_year = year(.$DATE)) %>% 
      # convert from monthly to yearly cpi values:
      group_by(cpi_year) %>% 
      dplyr::summarize(cpi = mean(VALUE)) %>% 
      # create adjustment factor relative 2010 year's prices:
      mutate(adj_fac2010 = cpi/cpi[cpi_year == 2010],
             adj_fac2015 = cpi/cpi[cpi_year == 2015])

# cpi from SOEP
cpi_raw_s <- read.table("/Users/LR/Git/clone/soeples/data/cpi.csv", header = T)
cpi_s <- cpi_raw_s %>% 
      select(year, cpi_west, cpi_ost) %>%
      # add missing values based on destatis data
      add_row(year = c(2014,2015), 
              cpi_west = c(106.6, 106.9),
              cpi_ost = c(106.6, 106.9)) %>% 
      mutate(cpi_ost = as.numeric(as.character(cpi_ost)),
             adj_fac2010_s = cpi_west / cpi_west[year == 2010],
             adj_fac2015_s = cpi_west / cpi_west[year == 2015]
      )

```

```{r}
#  merge datasets
mydf_raw <- merge(x = ppfadl, y = pgen, by= c("pid", "syear"), all.x = T) %>% 
            merge(., cpi, by.x = c("syear"), by.y=c("cpi_year"), all.x = T) %>% 
            merge(., cpi_s, by.x = c("syear"), by.y=c("year"), all.x = T) %>% 
      mutate(cpi_s = case_when(
            sampreg == 2 ~ cpi_ost, 
            sampreg == 1 ~ cpi_west)) %>% 
      select(-cpi_west, -cpi_ost)
```

```{r }
#### 2. cleaning data ####
#  2.1 Define Missings
      na_codes <- c(-1: -8)
      for (i in seq_along(mydf_raw)) {
            mydf_raw[[i]][mydf_raw[[i]] %in% na_codes] <- NA
      }
```


```{r}
# 2.2 Generate Variables
ex1 <- mydf_raw %>% 
      # Demographics
      mutate(ost = sampreg - 1, # ost = 1 west = 2
      # Einkommen, Hourly Wage
             breink10 = pglabgro /cpi_s*100
            )

      # Erwerbsstatus
      ex1$erwstatus <- ex1 %>% 
      mutate(erwstatus = case_when(
                              pgemplst == 1 ~ 1, 
                              pgemplst %in% c(2, 4) ~ 2
                              )
            ) %>%
      select(erwstatus) %>% unlist() %>% 
      labelled(c("Vollzeit" = 1, "Teilzeit" = 2)) 
      
#### 3. Describe Data ####
# Means by Region 
ergebnis1 <- ex1 %>% 
      filter(pop <= 2) %>% 
      select(syear, breink10, erwstatus, ost, phrf) %>% 
      dplyr::group_by(syear, ost) %>% 
      # mutate_at(c("breink10"), funs(Total = mean(., na.rm = T))) %>% 
      dplyr::summarise(TotalMean = mean(breink10, weights = phrf, na.rm = T))

     xtabs(TotalMean ~ syear + ost, ergebnis1) 

# Means by Region and Job Status
ergebnis2 <- ex1%>% 
      group_by(syear, ost, erwstatus) %>%
      dplyr::summarise(mean = mean(breink10, weights = phrf, na.rm = T))

xtabs(mean ~ syear + erwstatus + ost, ergebnis2) 

# ergebnis3 <- ex1 %>% 
#       group_by(syear, ost, erwstatus) %>%
#       select()
      
# to do merge so that totals are also in output
# ergebnis3 <- merge(ergebnis2, ergebnis1, by=c("syear", "ost"), incomparables = NA)
# try addmargins()
# try this https://stackoverflow.com/questions/33565949/add-row-to-data-frame-with-dplyr
# also try add_tally()
# for ergebnis3 http://www.stats.uwo.ca/faculty/murdoch/ism2013/4tables.pdf

     
# save dataset
#### save asample ####
saveRDS(ex1, file = "../../data/ex1.rds")
```

