---
title: ""
author: Lisa Reiber
output: 
  html_notebook: 
    code_folding: hide
    toc: yes
    toc_float: true
---

# Exercise 3

# Load Packages and Data
```{r}
#options(scipen = 100)
options(digits = 2)

#### install missing packages and load needed packages ####
if (!require("pacman")) install.packages("pacman")
pacman::p_load(foreign, readstata13, ggplot2, lubridate, broom, effects, 
               pastecs, interplot, ggeffects, dplyr)      

#### load dataset ####
load(file = "../../data/ex2.RData")
```


```{r}
## drop unplausible cases, generate sample for analyses
asample <- ex2 %>% 
        filter(pgtatzeit > 5,
               alter %>% dplyr::between(18, 64),
               pgemplst %in% c(1,2,4)
               ) %>% 
      # Unplausible Einzelfälle
      mutate(na = case_when(
              pid %in% c(1380202, 1380202, 607602, 2555301) ~ 1,
              pid == 8267202 & syear == 2007 ~ 1,
              pid == 2633801 & syear == 2006 ~ 1,
              pid == 2582901 & syear > 2006 ~ 1 )
             ) %>% 
        filter(is.na(na))

## Sample For Analysis for 2015
asample15 <- asample %>% filter(syear == 2015)

#### save mydf from ex2 ####
# saveRDS(ex3, file = "../../data/ex3.rds" )
# no need to save because no new variables generated
```

## 3.1 
You should start with an income equation with hourly gross wages (do not use the log-transfromation for now). Predictor variables are years of education, labor force experience, sex and region.
```{r}
(fit1  <- asample15 %>% 
            lm(hwageb~ pgbilzeit + erf + ost + frau, data=., weights=phrf))
      summary(fit1)
```
## 3.2 Specify an interaction effect for sex and years of education. Interpret the results. 
```{r}
# fit2 = add interaktion
(fit2 <- asample15 %>% 
      lm(hwageb ~ pgbilzeit*frau + erf + ost, data= ., weights = phrf))

summary(fit2)
```

## 3.3 Present your results grafically using a marginal effects plot
```{r}
# m = the object of a regression result
      # var1 = var whose coefficient is to be plotted 
      # var2 = var on which the coefficient is conditional (var2)
      
      # if we intend to know how the year of education can affect the coefficient for gender on the hourly wage, var1 is gender, and var2 is years of education. 
      
      # parameters:   point = wether a line or points are shown
      #               hist = it can be helpful to the evaluation of the substantive significance of the conditional effect to know the distribution of the conditioning variable (years of education)

interplot(m = fit2 , var1 = "frau", var2= "pgbilzeit", point = F, hist = T) +
            theme(axis.text.x  = element_text(angle=45)) +
            xlab("Education (in years)") +
            ylab("Estimated Coefficient for\nGender") +
            ggtitle("Estimated Coefficient of Gender \n on Hourly Wage by Years of Education") +
            theme(plot.title = element_text(face="bold")) +
            # Add a horizontal line at y = 0
            geom_hline(yintercept = 0, linetype = "dashed")

 # -> negative slope, shows that with increasing years of education(along the x axis), the magnitude of the coefficient of gender on the mileage decreases (along the y axis).
      
# Similarly, to show how the gender affects the coefficient of years of education on hourly wage, one only needs to switch var1 and var2:
      
interplot(m = fit2 ,var1 = "pgbilzeit" , var2="frau")

# margins paket, passt margins von Stata an
```

## 3.4 Specify another interaction effect between years of education and labor force experience. Interpret the results. Use a marginal effects plot to do that.

```{r}
 (fit4 <- sample15 %>% 
                  lm(hwageb ~ pgbilzeit*erf + frau + ost, data = ., weights = phrf))
      
      summary(fit4)
      
  # Marginal Effects Plot
      # Marginal Effects of Work Experience
      interplot(m=fit4, var1 = "pgbilzeit", var2="erf", point = F, hist = T) +
      theme(axis.text.x  = element_text(angle=45)) +
      xlab("Education (in years)") +
      ylab("Estimated Coefficient for\n Work Experience") +
      ggtitle("Estimated Coefficient of Work Experience \n on Hourly Wage by Years of Education") +
      theme(plot.title = element_text(face="bold")) +
      # Add a horizontal line at y = 0
      geom_hline(yintercept = 0, linetype = "dashed")
      
      # Marginal Effect of Years of Education
      interplot(m = fit4 ,var1 = "erf", var2 = "pgbilzeit", hist = T)  +
            xlab("Education (in years)") +
            ylab("Estimated Coefficient for\n Years of Education") +
            ggtitle("Estimated Coefficient of Years of Education \n on Hourly Wage by Labor Force Experience") +
            theme(plot.title = element_text(face="bold")) +
            geom_hline(yintercept = 0, linetype = "dashed")

      # Three way Interaction erf, pgbilzeit, frau
      
      dat <- ggpredict(fit4, terms = c("erf", "pgbilzeit", "frau"))

      ggplot(dat, aes(x, predicted)) +
            geom_line() +
            geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1)
      
      ggplot(dat, aes(x, predicted, colour = group)) +
            geom_line() +
            labs(
                  colour = get_legend_title(dat),
                  x = get_x_title(dat),
                  y = get_y_title(dat),
                  title = get_title(dat)
            )  +
            scale_color_manual(
                  values = c("red", "green", "blue"),
                  labels = get_legend_labels(dat)
            )
```
## 3.5 Include a nominal (e. g. company size: pgallbet) variable with more than two categories as predictors using a dummy set.
```{r}
(fit5 <- sample15 %>% 
                  lm(hwageb ~ pgbilzeit*frau + 
                              pgbilzeit*erf + 
                              ost + 
                              factor(pgallbet), 
                     data=., weights=phrf))
      
summary(fit5)
```

## 3.6 Drop the interactions and the dummy set. Include duration of unemployment as additional predictor. What has the largest effect, years of education, labor force experience or duration of unemployment
```{r}
(fit6 <- sample15 %>% 
            lm(hwageb ~ pgbilzeit + 
                        erf + 
                        frau + 
                        ost + 
                        pgexpue, 
               data=., weights=phrf))
      
      summary(fit6)
      
      anova(fit6)
      # years of education have larges partial effect size

# lm.beta package
# Doc: https://www.rdocumentation.org/packages/lm.beta/versions/1.5-1/topics/lm.beta
     lm.fit6.beta <- lm.beta(fit6)
     print(lm.fit6.beta)
     summary(lm.fit6.beta)
     coef(lm.fit6.beta)
```

## 3.7 Specify a quadratic effect for labor force participation. Interpret the result using a marginal effects plot.
```{r}
 # I(e17age^2)  
    # mutate(erf2 = erf^2) %>% 
           


      (fit7 <- sample15 %>%
                  lm(hwageb ~ pgbilzeit + 
                           erf + I(erf^2) + 
                           frau + 
                           ost + 
                           pgexpue, 
                     data=., weights=phrf))
      
      summary(fit7)

      # interplot(m = fit7 ,var1 = "erf", var2= "erf2", hist = T)  +
      #       xlab("Education (in years)") +
      #       ylab("Estimated Coefficient for\n Years of Education") +
      #       ggtitle("Estimated Coefficient of Years of Education \n on Hourly Wage by Labor Force Experience") +
      #       theme(plot.title = element_text(face="bold")) +
      #       # Add a horizontal line at y = 0
      #       geom_hline(yintercept = 0, linetype = "dashed")
      # 
      
      # ggpredict(fit7, terms = c("erf", "erf2"))
      
      # library(ggplot2)
      # ggplot(ggpredict(fit7, terms = c("erf")), aes(x, predicted, colour = group)) +
      #       geom_line()
      
      # dat <- ggpredict(fit7, "erf")
      
      
 # for polynomial data
      dat <- ggpoly(fit7, "erf")

      ggplot(dat, aes(x, predicted)) +
            stat_smooth(se = FALSE) +
            geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .15) +
            labs(x = get_x_title(dat), y = get_y_title(dat))
```