---
author: Lisa Reiber
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    code_folding: show
    toc: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
```

# Load Packages and Data
```{r}
#options(scipen = 100)
options(digits = 2)

#### install missing packages and load needed packages ####
if (!require("pacman")) install.packages("pacman")
pacman::p_load(foreign, readstata13, haven, ggplot2, lubridate, broom, effects, 
               pastecs, interplot, ggeffects, rstudioapi, tidyr, dplyr)      
 # "plm",  # panel regression analysis
              # "memisc",
              # "stargazer",  # output of regression results in Latex tables
              # "tidyverse" # incl. the packages imp. for every day data analyses


#### load dataset ####
load(file = "../../data/ex2.RData")
```

```{r}
## drop unplausible cases, generate sample for analyses
asample <- ex2 %>% 
        filter(pgtatzeit > 5,
               alter %>% dplyr::between(18, 64),
               pgemplst %in% c(1,2,4)
               ) %>% 
      # Unplausible Einzelfälle
      mutate(na = case_when(
              pid %in% c(1380202, 1380202, 607602, 2555301) ~ 1,
              pid == 8267202 & syear == 2007 ~ 1,
              pid == 2633801 & syear == 2006 ~ 1,
              pid == 2582901 & syear > 2006 ~ 1 )
             ) %>% 
        filter(is.na(na))

## Sample For Analysis for 2015
asample15 <- asample %>% filter(syear == 2015)

# prep: Ohne Anstaltsbevölkerung, Arbeitslose, Azubis, Rentner
# prep: ohne NA's in verwendeten Variablen
ex4 <- asample %>% 
      filter(pop < 3, 
             pgstib !=10, pgstib!=11, pgstib!=12, hwageb>0,
             phrf >0,
             syear %>% between(1991, 2015)
             ) %>% 
      dplyr::select(hwageb, pgbilzeit, pgisco88, pgnace, pgexpue,
             sex, ost, erf, pgallbet, phrf, syear, pid) %>%
      mutate(lnwage = log(hwageb),
             frau = sex-1) %>% 
      drop_na() 

#### save mydf from ex2 ####
saveRDS(ex4, file = "../../data/ex4.rds" )
```

# Exercise 4

## 4.1 Mincer Equation
Estimate the Mincer equation of last weeks exercise seperately for East and West Germany for all years since 1992 using loops (see foreach).

### Trend by Year
```{r trend}
# Fit Hourly Wage Brutto Model by year
      fit4 <- ex4 %>% 
            group_by(syear) %>% 
            do(model = lm(lnwage ~ 0+ pgbilzeit + 
                     erf + I(erf^2) + pgexpue +
                     ost + frau + factor(pgallbet) + factor(pgnace) +
                           factor(pgisco88), 
               data=., weights=phrf))

      # grab relevant info
      td_fit <- fit4 %>% tidy(model, conf.int = TRUE)
      td_plot <- td_fit %>% filter(term == "frau") 
      
      # Plot Over Years
      td_plot %>% 
            ggplot(aes(x = syear, y = estimate)) +
            geom_point() +
            geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                          alpha=0.3, color = "black") +
            geom_line() +
            ylab("Gender Revenue") +
            xlab("Survey Year") +
            ggtitle("Gender Pay Gap From 1992 to 2015") +
            scale_x_continuous(breaks = seq(1992, 2015, by= 5))+
            scale_y_reverse(labels = scales::percent)+
            theme_bw()
```

### Trend by Region
```{r Trend by Region}
#### Trend by Region ####
      fit4_region <- ex4 %>% 
            group_by(ost, syear) %>% 
            do(model = lm(lnwage ~ 0+ pgbilzeit + 
                     erf + I(erf^2) + pgexpue +
                     ost + frau + factor(pgallbet) + factor(pgnace) +
                           factor(pgisco88), 
               data=., weights=phrf))

      # look at model
            # (fit4_region$model) # only use if not too many models
            #fit4_region %>% tidy(model)
            #fit4_region %>% glance(model)
            #fit4_region %>% augment(model)
            
      # make tidy dataframe 
      (td_fit <- fit4_region %>% tidy(model, conf.int = TRUE))
      # select relevant info (coefficient for gender)
      td_plot <- td_fit %>% filter(term == "frau")
      
      # plot Trend by Region (grouped)
      td_plot %>% 
      ggplot(aes(x = syear, y = estimate, color = as.factor(ost))) +
            geom_point() +
            geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                          alpha=0.3, color = "black") +
            geom_line() +
            ylab("Gender Revenue") +
            xlab("Survey Year") +
            ggtitle("Gender Pay Gap by Region") +
            scale_x_continuous(breaks = seq(1991, 2015, by= 5))+
            scale_y_reverse(labels = scales::percent)
     
      # plot Trend by Region (facet)
      td_plot %>% 
      ggplot(aes(x = syear, y = estimate)) +
            geom_point() +
            geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                          alpha=0.3, color = "black") +
            geom_line() +
            facet_wrap(~ost)+
            ylab("Gender Revenue") +
            xlab("Survey Year") +
            ggtitle("Gender Pay Gap by Region") +
            scale_x_continuous(breaks = seq(1991, 2015, by= 5))
      
```

## 4.2 
Present the effects of years of education and of sex (the gender pay gap) graphically showing the trend since 1992 (see collapse) Tip: Use the Stata commands foreach, collapse and graph combine. 

```{r}

```

same but you use coefficient of frau rather than pgbilzeit

## 4.3 Tell Stata your data is panel data using xtset. Describe the variables of your Mincer equation and differentiate within and between variation (xtsum).

```{r}

```

