---
title: "ex8_soep"
author: "Lisa Reiber"
date: "10 1 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r, results=F, message=F, echo=FALSE}
# packages that are needed can be added underneath
p_needed <- c("readstata13",
              "tidyr",
              "strucchange",
              "plm",
              "pglm",
              "sandwich",
              "stargazer",
              "broom",
              "lfe",
              "lme4",
              "dplyr"
              )

packages <- rownames(installed.packages())
p_to_install <- p_needed[!(p_needed %in% packages)]
if (length(p_to_install) > 0) {
      install.packages(p_to_install)
      }
lapply(p_needed, require, character.only = TRUE)
```

```{r}
# Paths
inpath <- "/Volumes/gsoep32l/Gsoep32long/"     # home
path <- "~/Git/clone/panelregression2017"     # home

load(file = paste0(path,"/data/reg_sample.RData" ))
```

Open the healtl-file
```{r}
health_raw <- read.dta13("../../data/healthl.dta", convert.factors = F)
```

Does a divorce (famstat==4) causally affect mental health (mcs)?

1. Construct binary variables for the categories of famstat and a binary indicator of low mental health (mcs01 = 1 if mcs<40, otherwise 0)
```{r}

# Generate Binary Variables
health <- health_raw %>% 
      # mutate mcs
      mutate(mcs01 = ifelse(mcs<40, 1, 0)) %>% 
      # generate multiple binary vars from categorical var
      tibble::rownames_to_column() %>%
      group_by_all() %>% 
      summarise(count = n()) %>% 
      spread(famstat, count, sep = "_", fill = 0) %>% 
      ungroup() %>% 
      select(-rowname) %>% 
      drop_na()


health %>% 
      select(mcs, mcs01, starts_with("famstat")) %>% 
      head()
```

2. Estimate the effect of a divorce (fam4) on mental health (mcs01) in two different model specifications: Pooled Logit (logit, cluster()) and FE-Logit (xtlogit, fe).
```{r}
# Pooled Logit Model
plogit <- glm(mcs01 ~  famstat_4, family = binomial(), data = health)

# Fixed Effects Logit Model
felogit <- felm(mcs01 ~ famstat_4 | id , data = health)

felogit2 <- glm(mcs01 ~ famstat_4 + factor(id)-1, data = p.health,
                family= binomial)

# pglm package is not supported for paneldata
https://stats.stackexchange.com/questions/146434/why-pglm-fails-for-within-model

tidy(plogit)
tidy(felogit)

summary(plogit)
summary(felogit)

```

3. Add relevant third variables, if necessary.
```{r}
# OLS Model
plogit3 <- glm(mcs01 ~  famstat_4 + partner*gender + factor(occupation)+ hhinc, family = binomial, data = health)

# Fixed Effects Logit Model
p.health <- pdata.frame(health, index = c("id", "year"))
felogit3 <- plm(mcs01 ~  famstat_4 + partner*gender + factor(occupation) + hhinc, model = "within", data = p.health)

```

4. Summarize the estimates in a table (est tab)
```{r}
summary(plogit3)
summary(felogit3)

```

5. Interpret your findings and the differences between the models.

