---
author: "Lisa Reiber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook:
    code_folding: show
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Packages and Data
```{r, results=F, message=F, echo=FALSE}
#### install missing packages and load needed packages ####
if (!require("pacman")) install.packages("pacman")
pacman::p_load(foreign, readstata13, ggplot2, gtable, lme4, broom, memisc, tidyverse)      

#### load dataset ####
ess50 <- read.dta13(file = "../../data/ess50.dta", convert.factor = F)
```

# Exercise II
***
- Estimate a random intercept model using satisfaction with democracy as the outcome variable (
stfdem) and/or your previously selected outcomes.
- Generate a categorical variable (new) distinguishing between old/new democracies (democratization before 1950: country 1 2 3 5 6 8 9 10 13 14 15 16 17 18 21, between 1950 and 1989: country 7 20 11, and since 1989: country 4 12 19 22).
- Replicate the comparison of OLS versus random intercept models, regressing third variables on satisfaction with democracy (female and new).

**Data Wrangling**
```{r}
sample <- ess50 %>% 
      dplyr::select(cntry, stfdem, female, hinctnt, rlgdgr) %>% 
      mutate(cntry.id = as.numeric(as.factor(cntry)),
             new = ifelse(cntry.id %in% c(7,20,11),1,0),
             new2 = ifelse(cntry.id %in% c(4,12,19,22), 1, 0),
             old = ifelse(new == 0 & new2 == 1 , 1, 0)
             )
```


**OLS**
```{r}
# Regressions -------------------------------------------------------------

#OLS
fit.ols <- lm(stfdem ~ new + new2 + female, data = sample)
tidy(fit.ols)
```
**Random Intercept**
```{r}
# OLS with Random Intercept
fit.multi <- lmer(stfdem ~ new + new2 + female + (1|cntry.id), data = sample)
tidy(fit.multi)
```


# Exercise III
***

- Estimate a random intercept model using satisfaction with democracy as the outcome variable (stfdem). 
- Use one predictor variable with little distributional variation across countries (such as gender) and one predictor variable with substantial distributional variation across countries.
- Replicate the comparison of OLS versus random intercept models, regressing these predictor variables on satisfaction with democracy.
- Inspect standard errors and residual standard deviations at level 1 and 2 in particular. What explains the patterns of results you observe?

**Random Intercept: Little Variation across Countries**
```{r}
# Random Intercept
# Little Variation across countries
fit.multi.low <- lmer(stfdem ~ female + (1|cntry.id), data = sample)
tidy(fit.multi.low)
```
**Random Intercept: Substantial Variation across countries**
```{r}
# Random Intercept
# Substantial distrubutional variation across countries (income)
fit.multi.high <- lmer(stfdem ~ hinctnt + (1|cntry.id), data = sample)
tidy(fit.multi.high)

```

**OLS: Little Variation across countries**
```{r}
# OLS
# Little Variation across countries
fit.ols.low <- lm(stfdem ~ female, data = sample)
tidy(fit.ols.low)
```

**OLS: Substantial Variation across countries**
```{r}
# OLS
# Substantial distrubutional variation across countries (income)
fit.ols.high <- lm(stfdem ~ hinctnt, data = sample)
tidy(fit.ols.high)
```

**Standard Errors**
```{r, comment=NA }
## SE
# sqrt(diag(vcov(fit.ols.low))) OR tidy(fit.ols.low)

## RSD
#sigma(fit.multi.high) OR glance(fit.ols.low)

# (mtable_multi <- mtable("Low Var Model"=fit.multi.low,"High Var Model"=fit.multi.high,
#      summary.stats=c("sigma","R-squared","F","p","N")))
# 
# (mtable_ols <- mtable("Low Var Model"=fit.ols.low,"Low Var Model"= fit.ols.high,
#      summary.stats=c("sigma","R-squared","F","p","N")))


mtable(fit.ols.low, fit.ols.high)
mtable(fit.multi.low, fit.multi.high)

# xtable(anova(fm2b, fm2))
```
**Residual Standard Deviations at Level 1 and 2**
```{r}
glance(fit.ols.high)
glance(fit.ols.low)
glance(fit.multi.high)
glance(fit.multi.low)
```


# Extra: Visualization

**Coefficients**
```{r}

td <- tidy(fit.ols.high, conf.int = TRUE)
g1 <- ggplot(td, aes(estimate, term, color = term)) +
    geom_point() +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
      ggtitle("High Variation Model")

td <- tidy(fit.ols.low, conf.int = TRUE)
g2 <- ggplot(td, aes(estimate, term, color = term)) +
    geom_point() +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
            ggtitle("Low Variation Model")


grid.arrange(g1,g2, nrow = 2) 
```

